Value-Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking Around

### 2017-09-25 paper/paper.tex
Al Gore rythmn

\documentclass{article}
\usepackage{ijcai17}
\usepackage{times}
\usepackage{microtype}
\usepackage{amsmath}

% \usepackage{showframe}

@@

@@

\begin{description}
    \item[Tailored] Where we select vertices from highest degree to lowest degree (which is done by
        permuting the graph at the top of search).
    \item[Anti-Tailored] Where we select vertices from lowest degree to highest degree.
    \item[Extra Tiebreaking] Where we select vertices from highest degree to lowest degree,
        tiebreaking on highest sum of neighbour degrees.
    \item[Random] Where we select a branching vertex at random by (dynamically, that is, we shuffle
        the vertices before entering the for loop, rather than by randomly permuting the graph once at the
        top of search).
    \item[Tailor-Weighted Random] Where we shuffle the vertices before each for loop, but instead of
        picking a vertex uniformly at random, we use a weight of
        $2^{\left|\operatorname{N}(v)\right|}$ when deciding whether to branch on vertex $v$
        next.\footnote{Actually, we cheat a bit to avoid range issues with random number generation:
        we give the vertex with highest degree in the entire graph weight $2^{50}$, and then give
        lower degree vertices smaller powers, giving every vertex weight at least 1. This is a bit
        of hack but getting it perfectly right is hard.} This scheme heavily prefer vertices of high
        degree, but will give the same chance of being selected next to two vertices of equal
        degree. This also introduces an element of randomness, which will be important for restarts.
\end{description}

>>>>

\documentclass{article}
\usepackage{ijcai17}
\usepackage{times}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{tikz}

\usepackage{showframe}

\newcommand{\neighbourhood}{\operatorname{N}}
\newcommand{\vertexset}{\operatorname{V}}
\newcommand{\nds}{\operatorname{S}}

\usetikzlibrary{decorations, decorations.pathreplacing, decorations.pathmorphing,
    calc, backgrounds, positioning, tikzmark, patterns, fit}

\definecolor{chromaygb1}{rgb}{1.0, 1.0, 0.878431}
\definecolor{chromaygb1b}{rgb}{0.8, 0.972549, 0.72549}
\definecolor{chromaygb2}{rgb}{0.560784,0.929412,0.564706}
\definecolor{chromaygb3}{rgb}{0.407843,0.776471,0.560784}
\definecolor{chromaygb4}{rgb}{0.227451,0.631373,0.552941}
\definecolor{chromaygb5}{rgb}{0.301961,0.431373,0.670588}
\definecolor{chromaygb6}{rgb}{0.317647,0.254902,0.588235}
\definecolor{chromaygb7}{rgb}{0.294118,0.0,0.509804}

@@

\newcommand{\siplabel}[1]{\label{line:sip:#1}}
\begin{algorithm}[p]\DontPrintSemicolon\small
    \begin{tikzpicture}[remember picture,overlay]
        \coordinate (restart1sc) at ($(pic cs:restart1s) + (0, 0.15)$);
        \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
        \node [fill=chromaygb3, rounded corners, fit=(restart1sc) (restart1ec)] { };
        \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
        \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
        \node [fill=chromaygb3, rounded corners, fit=(restart2sc) (restart2ec)] { };
        \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
        \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
        \node [fill=chromaygb3, rounded corners, fit=(restart3sc) (restart3ec)] { };
        \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
        \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
        \node [fill=chromaygb3, rounded corners, fit=(restart4sc) (restart4ec)] { };
        \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
        \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
        \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
        \node [fill=chromaygb3, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
        \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
        \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
        \node [fill=chromaygb3, rounded corners, fit=(restart6sc) (restart6ec)] { };
    \end{tikzpicture}

    \nl $\FuncSty{subgraphIsomorphism}$ (Graph $\mathcal{P}$, Graph $\mathcal{T}$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$\left|\vertexset(\mathcal{P})\right| >
    \left|\vertexset(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\siplabel{enough}}
    \nl Discard isolated vertices in $\mathcal{P}$\siplabel{isolated} \;
    \nl \KwSty{global} $L \gets \big[ (\mathcal{P},\ \mathcal{T}), \siplabel{supplemental}
        (\mathcal{P}^{1,2}, \mathcal{T}^{1,2}),
        (\mathcal{P}^{2,2},\ \mathcal{T}^{2,2}), $\\\hspace*{6em}$ (\mathcal{P}^{3,2},\ \mathcal{T}^{3,2}),
        (\mathcal{P}^{4,2},\ \mathcal{T}^{4,2}) \big]$ \;
    \nl \ForEach{$v \in \vertexset(\mathcal{P})$}{
        \nl $D_v \gets \vertexset(\mathcal{T})$\;
        \nl \ForEach{$(P,\,T) \in L$}{
            \nl $D_v \gets \{ w \in D_v : $ \siplabel{degreefiltering} \\\hspace*{4em}$v \sim_P v \Rightarrow w \sim_T w~\wedge$
            \\\hspace*{4em}$\nds_P(v) \preceq \nds_T(w)\}$ \;
        }
    }
    \nl \lIf{$\KwSty{not}~\FuncSty{propagate}(D)$}{$\KwSty{return}~\KwSty{false}$}
    \nl \While{$\KwSty{true}$}{
        \nl \tikzmark{restart1s}$\KwSty{global}~(R, R') \gets (0,\,\textnormal{the next Luby value})$\tikzmark{restart1e} \;
        \nl $S \gets \FuncSty{search}(\{ E \in D : \left|E\right| > 1 \}, \top)$ \;
        \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
        \nl \lElseIf{$S = \KwSty{false}$}{$\KwSty{return}~\KwSty{false}$}
        \nl \tikzmark{restart2s}Trigger nogoods containing only one clause\tikzmark{restart2e} \;
    }
}
\medskip
    \nl $\FuncSty{search}$ (Domains $D$, Decisions $B$) $\rightarrow$ Bool
    \KwSty{or} \KwSty{restart} \;
\nl \Begin{
    \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
    \nl $D_v \gets \textnormal{the smallest domain in}~D$ \siplabel{pickd} \;
    \nl \tikzmark{restart3s}$C \gets \emptyset$\tikzmark{restart3e} \;
    \nl \ForEach {$v' \in D_v~\textnormal{ordered by chosen heuristic}$\siplabel{parforeach}}{
        \nl \tikzmark{restart4s}$C \gets C \cup \{ v' \}$\tikzmark{restart4e} \;
        \nl $D' \gets \FuncSty{clone}(D)$ \;
        \nl $D'_v \gets \{ v' \} $ \;
        \nl \If{\FuncSty{propagate}(D')}{
            \nl $S \gets \FuncSty{search}(\{ E \in D' : \left|E\right| > 1 \},$ \\\hspace*{3em} $B
            \wedge (v \mapsto v'))$ \;
            \nl \lIf{$S = \KwSty{true}$}{$\KwSty{return}~\KwSty{true}$}
            \nl \tikzmark{restart5s}\ElseIf{$S = \KwSty{restart}$}{
                \nl Post nogoods $\{B \wedge (v \mapsto c) \rightarrow \bot : c \in C \}$\tikzmark{restart5r}\;
                \nl $\KwSty{return}~\KwSty{restart}$\tikzmark{restart5e} \;
            }
        }
    }
    \nl \tikzmark{restart6s}\lIf{$(R \gets R + 1) = R'$}{$\KwSty{return}~\KwSty{restart}\tikzmark{restart6e}$}
    \nl $\KwSty{return}~\KwSty{false}$
}
\medskip
\nl $\FuncSty{propagate}$ (Domains $D$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \While{$D_v \gets$~\textnormal{a unit domain from}~$D$}{
        \nl $v' \gets $ the single value in $D_v$ \;
        \nl Trigger nogoods containing $(v \mapsto v')$ \;
        \nl \ForEach{$D_w \in D - D_v$}{
            \nl $D_w \gets D_w - v'$ \siplabel{removev} \;
            \nl \ForEach{$(P,\,T) \in L$}{
                \nl \lIf{$v \sim_P w$}{$D_w \gets D_w \cap \neighbourhood_T(v')$
                }
            }
            \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
        }
    }
\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$ \;
\nl \ForEach{$D_v \in D$ \textnormal{from smallest cardinality upwards\siplabel{eachdomain}}}{
    \nl $D_v \gets D_v \setminus H$ \siplabel{elimhall} \;
    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \siplabel{acc} \;
    \nl \lIf{$D_v = \emptyset~\vee~|A| < n$}{$\KwSty{return}~\KwSty{false}$\siplabel{failhall}}
    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}
}
\nl $\KwSty{return}~\KwSty{true}$ \;
}
\caption{Subgraph isomorphism. Highlighted lines are ignored for the non-restarting algorithm.}
\end{algorithm}

@@

\begin{description}
    \item[Tailored] Where we select vertices from highest degree to lowest degree (which is done by
        permuting the graph at the top of search).
    \item[Anti-Tailored] Where we select vertices from lowest degree to highest degree.
    \item[Extra Tiebreaking] Where we select vertices from highest degree to lowest degree,
        tiebreaking on highest sum of neighbour degrees.
    \item[Random] Where we select a branching vertex at random by (dynamically, that is, we shuffle
        the vertices before entering the for loop, rather than by randomly permuting the graph once at the
        top of search).
    \item[Tailor-Weighted Random] Where we shuffle the vertices before each for loop, but instead of
        picking a vertex uniformly at random, we use a weight of
        $2^{\left|\neighbourhood(v)\right|}$ when deciding whether to branch on vertex $v$
        next.\footnote{Actually, we cheat a bit to avoid range issues with random number generation:
        we give the vertex with highest degree in the entire graph weight $2^{50}$, and then give
        lower degree vertices smaller powers, giving every vertex weight at least 1. This is a bit
        of hack but getting it perfectly right is hard.} This scheme heavily prefer vertices of high
        degree, but will give the same chance of being selected next to two vertices of equal
        degree. This also introduces an element of randomness, which will be important for restarts.
\end{description}
